\Large The psychological arrow of time drives temporal asymmetries in inferring unobserved past and future events

### 2023-09-13 paper/main.tex
integrated replication experiment up until description of meta analysis (not including meta analysis description)

\newcommand{\stimDescription}{S1} % table
\newcommand{\segments}{S1} % figure
\newcommand{\events}{S2} %
\newcommand{\references}{S3}
\newcommand{\MethodsReplExp}{S4}
\newcommand{\targetAsymmetries}{S5}
\newcommand{\hitRates}{S6}
\newcommand{\characterRefs}{S7}
\newcommand{\refAdjacent}{S8}

@@

\textbf{JRM STOPPED HERE...}

\begin{figure}[tp]
  \centering
  \includegraphics[width=\textwidth]{results5}
  \caption{\textbf{Referenced events are associated with higher hit rates, but referring events are not.}  \textbf{A. Illustration of annotation approach.}  We extended the annotation procedure depicted in Figure~\ref{fig:result3}A to also label which events \textit{contained} references to events in other segments.  \textbf{B. Referenced versus referring events.}  During event $i$, when a character makes a reference to another event ($j$), we define $i$ as the \textit{referring} event and $j$ as the \textit{referenced} event.  \textbf{C.  Referring rate as a function of lag.}  Across all possible just-watched segments (lag 0), the bar heights denote the across-segment mean proportions of events containing references to events in other past (orange, negative lags) or future (blue, positive lags) segments.  The bar colors are described in the Figure~\ref{fig:result2} caption.  \textbf{D. Hit rates and counts of referenced, referring, and other events.}  As a function of temporal distance to the just-watched segment, the sub-panels display the numbers ($x$-axes) and hit rates ($y$-axes) of referenced (red), referring (green), and other (gray) events that participants hit (darker shading) or missed (lighter shading) in their uncued retrodictions (top sub-panel) and uncued predictions (bottom sub-panel).}
  \label{fig:result5}
\end{figure}

The preceding analyses show that when characters reference past or future events, those referenced events, and other events that are temporally adjacent to the referenced events, are more likely to be retrodicted and predicted.  In other words, referring to a past or future event in conversation leads to a ``boost'' in that event's hit rate.  We wondered whether this boost was bi-directional.  In particular: when a character refers (during a \textit{referring event}) to another event (i.e., the \textit{referenced event}), does this boost only the referenced event's hit rate, or does the referring event also receive a boost?  We labeled each event as a ``referring event,'' a ``referenced event,'' or a ``other event'' (i.e., not referring or referenced; Fig.~\ref{fig:result5}A, B).  We limited our analysis to references to onscreen (explicit) events. Consistent with our analysis of the proportions of referenced events (Fig.~\ref{fig:result3}B), the proportions of \textit{referring} events exhibited a \textit{forward} temporal asymmetry (Fig.~\ref{fig:result5}C).  Controlling for absolute lag, we found that referring events were associated with lower hit rates than referenced events (uncued retrodiction: OR = 0.03, $Z = -4.81$, $p < 0.001$, CI: 0.01 to 0.11; uncued prediction: OR = 0.04, $Z = -5.84$, $p < 0.001$, CI: 0.01 to 0.12; Fig.~\ref{fig:result5}D) and had no reliable differences in hit rates compared with other events (uncued retrodiction: OR = 0.37, $Z = -1.46$, $p = 0.15$, CI: 0.10 to 1.41; uncued prediction: OR = 2.16, $Z = 1.68$, $p = 0.09$, CI: 0.88 to 5.30).  This indicates that only referenced events received a hit rate boost (relative to other events), suggesting that the retrodictive and predictive benefits of references are directed (i.e., asymmetric).

>>>>

\newcommand{\stimDescription}{S1} % table
\newcommand{\segments}{S1} % figure
\newcommand{\events}{S2} %
\newcommand{\references}{S3}
\newcommand{\MethodsReplExp}{S4}
\newcommand{\targetAsymmetries}{S5}
\newcommand{\hitRates}{S6}
\newcommand{\characterRefs}{S7}
\newcommand{\refAdjacent}{S8}
\newcommand{\referringReferenced}{S9}

@@

\begin{figure}[tp]
  \centering
  \includegraphics[width=\textwidth]{results5}
  \caption{\textbf{Referenced events are associated with higher hit rates, but referring events are not.}  \textbf{A. Illustration of annotation approach.}  We extended the annotation procedure depicted in Figure~\ref{fig:result3}A to also label which events in our main experiment's stimuli \textit{contained} references to events in other segments.  \textbf{B. Referenced versus referring events.}  During event $i$, when a character makes a reference to another event ($j$), we define $i$ as the \textit{referring} event and $j$ as the \textit{referenced} event.  \textbf{C.  Referring rate as a function of lag.}  Across all possible just-watched segments (lag 0), the bar heights denote the across-segment mean proportions of events containing references to events in other past (orange, negative lags) or future (blue, positive lags) segments.  The bar colors are described in the Figure~\ref{fig:result2} caption.  \textbf{D. Hit rates and counts of referenced, referring, and other events.}  As a function of temporal distance to the just-watched segment, the sub-panels display the numbers ($x$-axes) and hit rates ($y$-axes) of referenced (red), referring (green), and other (gray) events that participants hit (darker shading) or missed (lighter shading) in their uncued retrodictions (top sub-panel) and uncued predictions (bottom sub-panel). For a display of analogous results from our replication experiment see Figure~\referringReferenced.}
  \label{fig:result5}
\end{figure}

The preceding analyses show that when characters reference past or future events, those referenced events, and other events that are temporally adjacent to the referenced events, are more likely to be retrodicted and predicted.  In other words, referring to a past or future event in conversation leads to a ``boost'' in that event's hit rate.  We wondered whether this boost was bi-directional.  In particular: when a character refers (during a \textit{referring event}) to another event (i.e., the \textit{referenced event}), does this boost only the referenced event's hit rate, or does the referring event also receive a boost?  We labeled each event as a ``referring event,'' a ``referenced event,'' or a ``other event'' (i.e., not referring or referenced; Fig.~\ref{fig:result5}A, B).  We limited our analysis to references to onscreen (explicit) events. Consistent with our analysis of the proportions of referenced events (Fig.~\ref{fig:result3}B), the proportions of \textit{referring} events exhibited a \textit{forward} temporal asymmetry (Fig.~\ref{fig:result5}C).  Controlling for absolute lag, we found that referring events were associated with lower hit rates than referenced events in our main experiment (uncued retrodiction: OR = 0.03, $Z = -4.81$, $p < 0.001$, CI: 0.01 to 0.11; uncued prediction: OR = 0.04, $Z = -5.84$, $p < 0.001$, CI: 0.01 to 0.12; Fig.~\ref{fig:result5}D) and had no reliable differences in hit rates compared with other events (uncued retrodiction: OR = 0.37, $Z = -1.46$, $p = 0.15$, CI: 0.10 to 1.41; uncued prediction: OR = 2.16, $Z = 1.68$, $p = 0.09$, CI: 0.88 to 5.30).  We also observed this phenomenon in our replication experiment (referenced events, uncued retrodiction: OR = XXX, $Z = XXX$, $p = XXX$, CI: XXX to XXX; referenced events, uncued prediction: OR = XXX, $Z = XXX$, $p = XXX$, CI: XXX to XXX; other events, uncued retrodiction: OR = XXX, $Z = XXX$, $p = XXX$, CI: XXX to XXX; other events, uncued prediction: OR = XXX, $Z = XXX$, $p = XXX$, CI: XXX to XXX;Fig.~\referringReferenced).  Taken together, this indicates that only referenced events received a hit rate boost (relative to other events), suggesting that the retrodictive and predictive benefits of references are directed (i.e., asymmetric).

\textbf{JRM STOPPED HERE}

\begin{figure}[tp]
  \centering
  \includegraphics[width=\textwidth]{meta-analysis}
  \caption{\textbf{Meta analysis}}
  \label{fig:meta-analysis}
\end{figure}
